Analyzing X-ray images in Python with scikit-image

### 2016-08-21 skimage_tomo.tex
Sections Getting started and Limitations

Several software applications and libraries are available to synchrotron
users to process their images. ImageJ \citep{Abramoff2004, Schneider2012}
and its distribution Fiji \citep{Schindelin2012} is a popular tool for
2-D and 3-D images, thanks to its intuitive menus and graphical tools,
and the wealth of plugins contributed by a vivid
community \citep{Schindelin2015}. For 3-D images, commercial tools such
as Avizo 3D software (TM) are appreciated for an intuitive graphical
pipeline and advanced 3D visualization. Alternatively, the use of a
programming language gives finer control and more possibilities, provided
classical processing algorithms can be called from libraries -- thereby
limiting the complexity of the programming task and the risk of bugs. Matlab (TM) and its
image processing toolbox are popular in the academic community of
computer vision and image processing. The Python language is widely used
in the scientific world and in synchrotron facilities. As a
general-purpose language, Python is used in synchrotrons to control
device servers \citep{pytango}, to access raw data of X-ray
detectors \citep{Knudsen2013}, to reconstruct tomography volumes from
radiographs \citep{Gursoy2014, Mirone2014}, and in data
processing packages for macromolecular cristallography \citep{Adams2010},
azimuthal integration of diffraction data \citep{Ashiotis2015}, or
fluorescence analysis \citep{pymca}.

@@

A stack of 2-D images, such as tomography slices generated by a reconstruction algorithm, can be opened as an image collection or a 3-D array:
\lstinputlisting[language=Python]{io_coll.py}
Raw data formats can be opened using the \texttt{NumPy} functions \texttt{fromfile} or \texttt{memmap}, and \texttt{hdf5} files are accessed using modules such as \texttt{h5py}, \texttt{pytables}.

@@

The Jupyter notebook \citep{Kluyver2016} is a web application that grew
out of the IPython project. Jupyter notebooks provide an interactive development
environment within a web browser, where live code can be enriched
by explanatory text, equations and visualizations (Fig.~\ref{fig:jupyter}).
As of July 2016, more than 500,000 Jupyter notebooks were posted on GitHub,
demonstrating their wide adoption by the community as workflow-sharing
tools (\url{http://archive.ipython.org/media/SciPy2016JupyterLab.pdf}).

@@

@@

Given the large size of tomography datasets, the execution speed of image
processing operations is of critical concern. \texttt{scikit-image} relies mostly
on calls to \texttt{NumPy} operations, of which most are performed in
optimized compiled code (C or Fortran). Performance-critical parts of \texttt{scikit-image}
that cannot call efficient \texttt{NumPy} code are implemented in
\texttt{Cython}. \texttt{Cython} \citep{Behnel2011} is an extension of the
Python language that supports explicit type declarations, and is compiled
directly to C.

@@

An easy solution consists in dividing an image into smaller images (with
or without overlap, depending on the operation), and to apply the
same operation on the different sub-images, on different cores. Creating overlapping chunks is easy with the dedicated function \texttt{view\_as\_windows} (or \texttt{view\_as\_blocks} for contiguous non-overlapping chunks):
\begin{lstlisting}[language=Python]
>>> from skimage import util, data
>>> im = data.camera()
>>> im.shape
(512, 512)
>>> # chunks of size 120 with 8-pixel overlap
>>> chunks = util.view_as_windows(im, (128, 128), step=120)
>>> chunks.shape
(4, 4, 128, 128)
\end{lstlisting}
The \texttt{joblib} library enables easy parallel processing. Looping over the different blocks, and dispatching the computation over several cores, is realized with the following syntax:
\begin{lstlisting}
>>> from joblib import Parallel, delayed
>>> filtered_chunks = Parallel(n_jobs=4)(delayed(filters.gaussian)(chunks[i, j]) for i in range(4) for j in range(4))
\end{lstlisting}
\texttt{scikit-image} also offers experimental support for a more integrated parallel processing pipeline, thanks to the \texttt{dask} \citep{dask} module:
\begin{lstlisting}
filtered_im = util.apply_parallel(filters.gaussian, im, depth=8)
\end{lstlisting}
The size of chunks is determined automatically from the number of available cpus, or can be specified by the user.

@@

\section*{Documentation and getting help}

@@

>>>>

Several software applications and libraries are available to synchrotron
users to process their images. ImageJ \citep{Abramoff2004, Schneider2012}
and its distribution Fiji \citep{Schindelin2012} is a popular tool for
2-D and 3-D images, thanks to its intuitive menus and graphical tools,
and the wealth of plugins contributed by a vivid community
\citep{Schindelin2015}. For 3-D images, commercial tools such as Avizo 3D
software (TM) are appreciated for an intuitive graphical pipeline and
advanced 3D visualization. Alternatively, the use of a programming
language gives finer control, a better reproducibility and more
possibilities, provided classical processing algorithms can be called
from libraries -- thereby limiting the complexity of the programming task
and the risk of bugs. Matlab (TM) and its image processing toolbox are
popular in the academic community of computer vision and image
processing. The Python language is widely used in the scientific world
and in synchrotron facilities. As a general-purpose language, Python is
used in synchrotrons to control device servers \citep{pytango}, to access
raw data of X-ray detectors \citep{Knudsen2013}, to reconstruct
tomography volumes from radiographs \citep{Gursoy2014, Mirone2014}, and
in data processing packages for macromolecular cristallography
\citep{Adams2010}, azimuthal integration of diffraction data
\citep{Ashiotis2015}, or fluorescence analysis \citep{pymca}.

@@

A stack of 2-D images, such as tomography slices generated by a reconstruction algorithm, can be opened as an image collection or a 3-D array:
\lstinputlisting[language=Python]{io_coll.py}
Raw data formats can be opened using the \texttt{NumPy} functions
\texttt{fromfile} (to load the array into memory) or \texttt{memmap} (to
keep the array on disk). The following code creates an array from a
raw image file of unsigned 16-bit integers with a header of 1024 bytes 
\begin{lstlisting}
im = np.memmap('image.edf', shape=(2048, 2048), offset=1024, dtype=np.uint16)
\end{lstlisting}
For every raw data specification, it is thus very easy to write a reader
using \texttt{np.memmap} (see for example
\url{https://github.com/jni/python-redshirt}). 
\texttt{hdf5} files are accessed using modules such as \texttt{h5py}, \texttt{pytables}.
\texttt{hdf5} files are accessed using modules such as \texttt{h5py}, \texttt{pytables}.

@@

The Jupyter notebook \citep{Kluyver2016} is a web application that grew
out of the IPython project. Jupyter notebooks provide an interactive
development environment within a web browser, where live code can be
enriched by explanatory text, equations and visualizations
(Fig.~\ref{fig:jupyter}). Jupyter notebooks render directly as webpages
on GitHub, making them a straightforward tool to publish online a
script and its output. As of July 2016, more than 500,000 Jupyter
notebooks were posted on GitHub, demonstrating their wide adoption by the
community as workflow-sharing tools
(\url{http://archive.ipython.org/media/SciPy2016JupyterLab.pdf}).

@@

\paragraph{Capabilities.}

@@

Given the large size of tomography datasets, the execution speed of image
processing operations is of critical concern. \texttt{scikit-image}
relies mostly on calls to \texttt{NumPy} operations, of which most are
performed in optimized compiled code (C or Fortran). Performance-critical
parts of \texttt{scikit-image} that cannot call efficient \texttt{NumPy}
code are implemented in \texttt{Cython}. \texttt{Cython}
\citep{Behnel2011} is an extension of the Python language that supports
explicit type declarations, and is compiled directly to C. Therefore, the
performance of \texttt{scikit-image} can be close to the one of libraries
written in a compiled language such as C++ or Java. For example,
computing the watershed segmentation of a 2000x2000 array of floats into
1000 regions took about 1 s using 1 CPU of an off-the-shelf laptop, and
10 s for a 256x256x256 image segmented into 2000 regions. Similar
timescales were obtained with \texttt{mahotas}, a Python package
implemented exclusively in C++ \citep{Coelho2013} (with a slight advantage
for \texttt{mahotas}).

@@

An easy solution consists in dividing an image into smaller images (with
or without overlap, depending on the operation), and to apply the
same operation on the different sub-images, on different cores. Creating overlapping chunks is easy with the dedicated function \texttt{view\_as\_windows} (or \texttt{view\_as\_blocks} for contiguous non-overlapping chunks):
\begin{lstlisting}[language=Python]
>>> from skimage import util, data
>>> im = data.camera()
>>> im.shape
(512, 512)
>>> # chunks of size 134 with 8-pixel overlap
>>> chunks = util.view_as_windows(im, (134, 134), step=126)
>>> chunks.shape
(4, 4, 134, 134)
\end{lstlisting}
The \texttt{joblib} library enables easy parallel processing. Looping over the different blocks, and dispatching the computation over several cores, is realized with the following syntax:
\begin{lstlisting}
>>> from joblib import Parallel, delayed
>>> filtered_chunks = Parallel(n_jobs=4)(delayed(filters.gaussian)(chunks[i, j]) for i in range(4) for j in range(4))
\end{lstlisting}
\texttt{scikit-image} also offers experimental support for a more integrated parallel processing pipeline, thanks to the \texttt{dask} \citep{dask} module:
\begin{lstlisting}
filtered_im = util.apply_parallel(filters.gaussian, im, depth=8)
\end{lstlisting}
The size of chunks is determined automatically from the number of available cpus, or can be specified by the user.

@@

\section*{Documentation}

@@

\section*{Current limitations and challenges}

While have insisted on the assets of \texttt{scikit-image} for processing
X-ray images, one should be aware of current limitations of the package.

\paragraph{Speed of execution.} Although \texttt{scikit-image} approaches
the speed of execution of compiled (C++, Java) code, it cannot reach the
performance of code optimized for the GPU. \texttt{scikit-image} is not
the best tool for ultrafast computations where the workflow is known
beforehand, simple and stable. However, it is an excellent tool for
exploring image data interactively and testing different algorithms -- an
important component of data processing in scientific work --, and its
speed of execution is good enough to process the Gb-sized tomographic
images in seconds to minutes.

\paragraph{3-D compatibility.} At the moment, about two thirds of
\texttt{scikit-image} functions take indifferently 2D or 3D arrays. It is
planned to increase this compatibility in the future.

\paragraph{Documentation for domain-specific applications.} Some image
processing libraries or applications address a specific scientific
domains, such as CellProfiler \citep{Carpenter2006, Lamprecht2007} for
biological images. The documentation of such projects often showcases
examples that are close to the experience of the targeted community.
Since \texttt{scikit-image} is application-agnostic,
applications such as tomographic imaging are not mentioned in detail in the
documentation of \texttt{scikit-image}. A possible improvement would be
to write comprehensive tutorials addressing specific communities, and to
reference these tutorials from the main \texttt{scikit-image}
documentation.

\section*{Getting started}

Scientists interested in experimenting with \texttt{scikit-image} are
invited to read the installation instructions on
\url{http://scikit-image.org/docs/stable/install.html}.
\texttt{scikit-image} can be installed either bundled in a Scientific
Python distribution, such as Anaconda or Canopy, or alone (with its
dependencies) using a installer/packager program such as pip or Ubuntu's
Synaptic.

The Getting started section of the online User Guide
(\url{http://scikit-image.org/docs/dev/user_guide/getting_started.html})
is a good starting point, as well as the other sections of the user
guide. Examples of the gallery
(\url{http://scikit-image.org/docs/dev/auto_examples/}) can be used to
find applications close to one's needs. Although most examples of the
gallery use 2D images, most functions can be used with 3D images as well.

Assistance on matters not covered by the documentation is provided to
users either on the dedicated mailing-list
\url{scikit-image@googlegroups.com} or on StackOverflow
\url{http://stackexchange.com/filters/87964/scikit-image}.
